---
title: "Homework 3"
Author: Natalie Boychuk (nab2194)
output: github_document
---


```{r}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Problem 1

```{r}
data("instacart") 
```

This dataset contains `r nrow(instacart)` rows (representing the total number of observations) and `r ncol(instacart)` columns. 

Observations are the level of items in order by user. There are user/order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. FINISH UP THIS TEXT 

*How many aisles are there, and from which aisles are the most items ordered?* 

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are 134 aisles in this dataset. Among these, most items are ordered from the 'fresh vegetables,' 'fresh fruits,' and 'packaged vegetables fruits' aisles. Instacart shoppers are a healthy bunch! 

*Making a plot* 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

*Making a table with most popular products in baking ingredients, dog food care, and packaged vegetables/fruit* 
```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>% 
  knitr:: kable()
```

*Table with mean hour of day at which Pink Lady Apples/Coffee ice cream are ordered* 

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr :: kable()
```

## Problem 2 

```{r bullet point 1 - data tidying}
accel = 
  read_csv("./data/accel_data.csv") 

accel_longer = 
  pivot_longer(
  accel,
  activity.1:activity.1440,
  names_to = "minute_of_day",
  names_prefix = "activity.",
  values_to = "activity_count"
  ) 

accel_wkdy = 
mutate(accel_longer,
       weekday_end = case_when(
         day == "Monday" ~ "weekday",
         day == "Tuesday" ~ "weekday",
         day == "Wednesday" ~ "weekday",
         day == "Thursday" ~ "weekday",
         day == "Friday" ~ "weekday",
         day == "Saturday" ~ "weekend",
         day == "Sunday" ~ "weekend"
         )) %>% 
  mutate(weekday_end = as.factor(weekday_end)) %>% 
  relocate("week","day_id","day","weekday_end", "minute_of_day","activity_count") %>% 
  mutate(minute_of_day = as.double(minute_of_day)) %>% 
  mutate(day = as.factor(day)) %>%
  mutate(
  day = forcats::fct_relevel(day, "Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))

```

This dataset provides information on the activity counts for one man over every minute of a 5-week period. After tidying the data, the resulting dataframe has `r nrow(accel_wkdy)` rows and `r ncol(accel_wkdy)` columns. The dataframe includes the week number, an ID for the day number, a binary variable referring to whether the day is a weekday or a weekend (defined as Saturday/Sunday), the minute of the day, and the activity count.

```{r Making a table with aggregate activity}


  
```


Creating a plot 
```{r}

```



## Problem 3 

```{r loading in the dataset}


```

